Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-01-31 manuscript/climate_sensitivity_draft.Rmd
update related to issue #70
worked on intro

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought. @bennett_larger_2015 was the first to show greater growth reductions for larger trees on a global scale, and numerous subsequent studies have reinforced this finding (*e.g.*, @stovall_tree_2019; @hacket-pain_consistent_2016], although there are exceptions [@zuleta_drought-induced_2017]. However, it has yet to be resolved which of several potential underlying mechanisms drive this pattern. First, taller trees have a greater biophysical challenge of lifting water greater distances against the effects of gravity and friction [@mcdowell_relationships_2011; @mcdowell_darcys_2015; @ryan_hydraulic_2006; @couvreur_water_2018], especially during drought [@zhang_size-dependent_2009]. Second, larger trees may have lower drought resistance because of their higher crown positions, where they are exposed to higher solar radiation, greater wind speeds, and lower humidity [@koike_leaf_2001; @kunert_revised_2017 **REFS-KAT**). Subcanopy trees subsequently tend to fare better specifically due to the benefits of a buffered environment [@pretzsch_drought_2018]. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems confer a potential advantage in terms of allowing greater access to water; however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. 
*A final mechanism that could mediate tree size-related responses to drought is how hydraulic traits are distributed with respect to size [@meakem_role_2018; @liu_hydraulic_2019 [Roskilly et al. 2019](https://www.pnas.org/content/pnas/early/2019/06/14/1900734116.full.pdf)]. **{LS:—larger trees may have larger xylem conduits, which in turn may have less resistance to air seeding and subsequent embolism (Mark Olsen argument, right?) }** The pattern observed by @bennett_larger_2015 could be caused if the larger size classes were dominated by species less adapted to handle drought. Alternatively, larger size classes may be dominated by species that are inherently better adapted to greater biophysical challenges--as is the case in tropical moist forests of Panama, where larger size classes contain greater proportions of deciduous species [@condit_quantifying_2000; @meakem_role_2018]. Understanding the mechanisms driving the greater growth reductions of larger trees during drought will require sorting out the interactive effects of height, canopy position, soil water availability, and species' traits. * {see Lawren's comment on order of presentation}

@@

Our results partially support Hypothesis *H2.1*, where species' hydraulic traits --$XP$, $PLA_{dry}$, and $\pi_{tlp}$--were sometimes useful in explaining variation in drought responses, whereas $LMA$ and $WD$ never significantly associated with $Rt$ in the univariate models (all dAIC $\le$ 0.22; Tables 1,4,5). Therefore, these were excluded as candidate variables for the full multivariate models. In contrast, $XP$, $PLA_{dry}$, and $\pi_{tlp}$ all explained modest amounts of variation (dAIC > 1.0) in at least one drought (Table 4). $PLA_{dry}$ was a strong predictor for the 1966 drought and all droughts combined, with consistently negative coefficients. $\pi_{tlp}$ was not significant in the univariate models but had a consistently negative coefficient (Table 4). Whereas ring-porous species had highest $Rt$ overall and in the 1966 and 1999 droughts, they had lower $Rt$ in 1977. Results were similar in the context of multivariate models (Table 5), except that $PLA_{dry}$ had a small positive coefficient in the one 1999 model in which it was included. $PLA_{dry}$, and $\pi_{tlp}$ were each included in the top overall model and in the top model for two of the three individual droughts (Table 5).

>>>>

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought. @bennett_larger_2015 was the first to show greater growth reductions for larger trees on a global scale, and numerous subsequent studies have reinforced this finding (*e.g.*, @stovall_tree_2019; @hacket-pain_consistent_2016], although there are exceptions [@zuleta_drought-induced_2017]. However, it has yet to be resolved which of several potential underlying mechanisms drive this pattern. First, taller trees have a greater biophysical challenge of lifting water greater distances against the effects of gravity and friction [@mcdowell_relationships_2011; @mcdowell_darcys_2015; @ryan_hydraulic_2006; @couvreur_water_2018], especially during drought [@zhang_size-dependent_2009]. To meet this challenge, hydraulic traits vary in relation to tree height within indivdiuals {**REFS**}, within species {**REFS**}, and across species [@liu_hydraulic_2019]. Despite the tendency for hydrualic traits to become more conservative with height on an individual tree {**REFS**}, the need for greater hydraulic efficiency (*i.e.,* larger xylem conduits) in tall trees [@liu_hydraulic_2019]{**REFS**} may leave them more vulnerable during drought {**REFS-- Mark Olsen--mentioned by Lawren**}. Second, larger trees may have lower drought resistance because of their higher crown positions, where they are exposed to higher solar radiation, greater wind speeds, and lower humidity [@koike_leaf_2001; @kunert_revised_2017 **REFS-KAT**). Subcanopy trees tend to fare better specifically due to the benefits of a buffered environment [@pretzsch_drought_2018]. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems confer a potential advantage in terms of allowing greater access to water; however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. A final mechanism that could mediate tree size-related responses to drought is how species, and their associated hydraulic traits, are distributed with respect to size [@meakem_role_2018; @liu_hydraulic_2019]. Within a forest community, greater drought vulnerability of larger trees could occur if the larger size classes were dominated by species less adapted to handle drought. Alternatively, larger size classes may be dominated by species that are inherently better adapted to greater biophysical challenges--as is the case in tropical moist forests of Panama, where larger size classes contain greater proportions of deciduous species [@condit_quantifying_2000; @meakem_role_2018]. Understanding the mechanisms driving the greater growth reductions of larger trees during drought will require sorting out the interactive effects of height, canopy position, soil water availability, and species' traits.  {**see Lawren's comment on order of presentation**}

@@

Our results partially support Hypothesis *H2.1*, where species' hydraulic traits --$XP$, $PLA_{dry}$, and $\pi_{tlp}$--were sometimes useful in explaining variation in drought responses, whereas $LMA$ and $WD$ never significantly associated with $Rt$ in the univariate models (all dAIC $\le$ 0.22; Tables 1,4,5). Therefore, these were excluded as candidate variables for the full multivariate models. In contrast, $XP$, $PLA_{dry}$, and $\pi_{tlp}$ all explained modest amounts of variation (dAIC > 1.0) in at least one drought (Table 4). $PLA_{dry}$ was a strong predictor for the 1966 drought and all droughts combined, with consistently negative coefficients. $\pi_{tlp}$ was not significant in the univariate models but had a consistently negative coefficient (Table 4). Whereas ring-porous species had highest $Rt$ overall and in the 1966 and 1999 droughts, they had lower $Rt$ in 1977. Results were similar in the context of multivariate models (Table 5). $PLA_{dry}$, and $\pi_{tlp}$ were each included in the top overall model and in the top model for two of the three individual droughts (Table 5).
